# Simulation results 


## Survival times

```{r load-process-data}
dat <- readRDS('results/data-all.rds')

scen_levs <- c("simple", "noise", "noise_pred", "simple_int", "noise_int", "noise_pred_int")
labs_lng <- c("Basic", "Noise", "Predictive noise", "Interaction", "Interaction + noise", "Interaction + predictive noise")
labs_lng <- paste(labs_lng, "scenario")

dat <- dat %>% mutate(
  rep = (.data$ID - 1) %/% 200,
  mod = factor(
    .data$mod,
    levels = c('best', 'mars', 'rf', 'rpart', seq(0.1, 1, 0.1)),
    labels = c('Best', 'MARS', 'RF', 'CART', seq(0.1, 1, 0.1))
  ),
  scenario = factor(.data$scenario, levels = scen_levs, labels = labs_lng)
)

dat_summ <- dat %>%
  group_by(.data$mod, .data$samp, .data$scenario) %>%
  summarise(mean_surv_time = mean(exp(.data$tot_reward))) %>%
  ungroup()
```

```{r defn-plot-surv}
plot_surv <- function(dat) {
  ggplot(dat) +
    geom_boxplot(mapping = aes(x = mod, y = mean_surv_time)) +
    labs(y = "Mean months of survival", x = "Regime") +
    theme(legend.position = "none") +
    coord_trans(y = "log") +
    facet_wrap(~ scenario, nrow = 3) +
    theme_bw()
}
```

```{r results-no-int, fig.cap=caption, fig.scap='Mean months of survival under each regime for scenarios without interaction', fig.dim=c(7, 7)}
caption <- 'Mean months of survival under each regime for the scenarios without interaction, log scale. The mean months of survival on the test set (initial tumor masses and toxicities for 2000 patients) for the regimes estimated using Q-learning (MARS, RF--here equivalent to bagging, and CART), each of the 100 treatment regimes estimated from each of the 100 training sets are shown in box plots. The best regime and the constant dose regimes (\\(1, 0.9, \\ldots, 0.1\\)) do not vary across training set and therefore have only one mean survival time for the test set. Observations more than \\(1.5 \\times IQR\\) from either hinge (lower or upper quartile) are shown as points.'
plot_surv(dat_summ[!grepl('Interaction', dat_summ$scenario), ])
```

```{r results-int, fig.cap=caption, fig.scap='Mean months of survival under each regime for scenarios with interaction', fig.dim=c(7, 7)}
caption <- 'Mean months of survival under each regime for scenarios with interaction, log scale. As in Figure \\@ref(fig:results-no-int), mean months of survival for the regimes estimated using Q-learning (MARS, RF--here equivalent to bagging, and CART), each of the 100 treatment regimes estimated from each of the 100 training sets are shown in box plots. The best regime and the constant dose regimes (\\(1, 0.9, \\ldots, 0.1\\)) do not vary across training set and therefore have only one mean survival time for the test set. Observations more than \\(1.5 \\times IQR\\) from either hinge (lower or upper quartile) are shown as points.'
plot_surv(dat_summ[grepl('Interaction', dat_summ$scenario), ])
```

```{r rewards}
col_heads <- c("Regime", "B", "N", "PN", "I", "I+N", "I+PN")
cap <- "Mean months of survival under each scenario (in months), averaged over 2000 test patients for each regime. Each test set of 2000 had identical initial tumor masses and toxicities (\\(M_{0}\\), \\(W_{0}\\)). Best is the optimal dose sequence obtained from evaluating every possible three dose sequence. MARS, RF (random forest, which is equivalent to bagging in these scenarios), and CART (regression trees using the CART procedure) represent the average survival times when using the optimal dose regime estimated from Q-learning using each method to fit the Q-functions, averaged over the 100 training set replicates. B is the basic scenario, N is noise scenario, PN is the predictive noise scenario, I is the interaction scenario, I+N is the interaction and noise scenario, and I+PN is the interaction and predictive noise scenario. Each scenario is described in detail in section \\@ref(simulation)."
cap_toc <- "Mean survival times under each scenario"
dat %>%
  group_by(.data$scenario, .data$mod) %>%
  summarise(mean_surv_time = mean(exp(.data$tot_reward))) %>%
  spread(key = .data$scenario, value = .data$mean_surv_time) %>%
  knitr::kable(caption = cap, digits = 1, col.names = col_heads, booktabs = TRUE)
```

The arithmetic mean survival times under each treatment regime are shown in Figure \@ref(fig:results-no-int) for the scenarios without covariate by treatment interaction (patient subgroups with differential treatment response) and in Figure \@ref(fig:results-int) for the interaction scenarios--those with covariate by treatment interaction. "Best" shows the arithmetic mean survival time for the 2000 test patients when following the best possible dose sequence. The optimal regimes estimated using Q-learning are labelled as "MARS", "RF", and "CART" for Q-learning using MARS, random forest (which in these situations was equivalent to bagging), and regression trees using the CART procedure to estimate the Q-functions, respectively. In Figures \@ref(fig:results-no-int) and \@ref(fig:results-int), each of the 100 mean survival times obtained using the optimal regime derived from the 100 different training sets on the test set are displayed as separate points in box plots for the three modeling techniques. Comparison regimes are shown with only a single line because they do not require training and are therefore invariant across training set. Table \@ref(tab:rewards) displays the arithmetic mean survival times under each regime, averaged over all of the 100 test set survival times for each of the regimes estimated using each Q-learning technique on each training set.

```{r sd-train-rewards}
cap <- "Standard deviation of mean survival times across training set replicates for Q-learning models. B is the basic scenario, N is noise scenario, PN is the predictive noise scenario, I is the interaction scenario, I+N is the interaction and noise scenario, and I+PN is the interaction and predictive noise scenario. Each scenario is described in detail in section \\@ref(simulation)"
cap_toc <- "Standard deviation of mean survival times across training sets"
dat %>%
  filter(mod %in% c("BMARS", "MARS", "RF", "CART")) %>%
  group_by(scenario, mod, samp) %>%
  summarise(mean_surv_time = mean(exp(tot_reward))) %>%
  group_by(scenario, mod) %>%
  summarise(sd_mean_surv_time = sd(mean_surv_time)) %>%
  spread(key = scenario, value = sd_mean_surv_time) %>%
  knitr::kable(caption = cap, col.names = col_heads, digits = 1)
```

In the basic (B) scenario, MARS and random forest produce similar mean survival times, about two thirds of the best survival times, with MARS performing slightly better on average. However, as will be repeated throughout scenarios, MARS exhibits more sensitivity to the training set (higher variance) as shown in Figure \@ref(fig:results-no-int) and in Table \@ref(tab:sd-train-rewards), which shows standard deviations in mean months of survival across training sets. As is also repeated across scenarios, a single regression tree (labelled "CART" throughout) performs very poorly, worse overall than many constant dose regimes. When the 100 noise variables are added in the noise (N) scenario, there is approximately a 10 month decrease in mean survival times for MARS and RF, bringing them to about half of the best possible survival times and about a 2 month decrease for CART. MARS then performs slightly worse than random forest when noise variables are added. Allowing some of the noise variables to shift survival times up or down in the predictive noise (PN) scenario produces results very similar to the noise scenario without predictive noise.

Results from the interaction scenarios, scenarios with the four treatment subgroups created by \(X_{1}\) and \(X_{2}\), shown in Figure \@ref(fig:results-int) and Table \@ref(tab:rewards), were similar to those in the scenarios without interaction, but shifted towards longer survival. Despite the shift upwards, the trend is very similar to the scenarios without interaction. In the interaction only scenario (I), MARS and random forest produce similar mean survival times, a little more than half of those produced by the best possible regime, with MARS showing greater variability across training sets. Addition of noise variables to the interaction scenario (I+N), results in a reduction of around 10 months of survival time for MARS and random forest, bringing them to half or a bit less than half of the best possible survival time. As in the scenarios without interaction, MARS decreases more than random forest when the noise variable are added. Allowing 10 of the noise variables to shift survival up or down by a little in the interaction and predictive noise (I+PN) scenario has little effect on the performance of the estimated regimes. CART produces results very similar to those in the scenarios without interaction.


## Variable importances

```{r load-imps-data}
imps <- readRDS('results/data-importance.rds')
imps$scenario <- factor(imps$scenario,
                        levels = gsub('_', '-', scen_levs),
                        labels = labs_lng)
imps$var_grp <- replace(imps$var_nm, startsWith(imps$var_nm, 'V'), 'V')
imps$var_grp <- replace(imps$var_grp, startsWith(imps$var_nm, 'Z'), 'Z')
imps$var_grp <- replace(imps$var_grp, startsWith(imps$var_nm, 'X'), 'X')
imps$var_grp <- factor(imps$var_grp,
                       levels = c("tumor_mass", "toxicity", "dose", "X", "Z", "V"),
                       labels = c("M", "W", "D", "X", "Z", "V"))
imps$mod = factor(imps$mod, levels = c("mars", "rf", "rpart"),
                  labels = c("MARS", "RF", "CART"))
```

```{r imps-plot}
plot_imps <- function(data) {
    ggplot(data) +
    geom_violin(aes(x = var_grp, y = importance,
                    linetype = mod, color = mod),
                position = "dodge") +
    labs(x = "Variable",
         y = "Importance (% of most important variable)",
         linetype = "Model", color = "Model") +
    facet_wrap( ~ scenario, nrow = 3) +
    theme_bw() + theme(legend.position = "bottom")
}
```

```{r var-imps-no-int, fig.cap = cap, fig.scap='Variable importances for the scenarios without interaction', fig.dim=c(7, 7)}
cap <- "Variable importances for the scenarios without interaction. Importances are given as a percentage of the most important variable in each model. Importances from each stage and each training set replicate are shown together in each violin (e.g. each of the 60 importances for tumor mass for each stage for each model are shown together in the three respective violins above M for each scenario). All of the importances for each noise variable in the \\(Z\\) vector are labelled Z, while the variables in the \\(V\\) and \\(X\\) vectors are labelled V and X respectively."
plot_imps(imps[!grepl('Interaction', scenario), ])
```

```{r var-imps-int, fig.cap = cap, fig.scap='Variable importances for interaction scenarios', fig.dim=c(7, 7)}
cap <- "Variable importances for the interaction scenarios. Importances are given as a percentage of the most important variable in each model. Importances from each stage and each training set replicate are shown together in each violin (e.g. each of the 60 importances for tumor mass for each stage for each model are shown together in the three respective violins above M for each scenario). All of the importances for each noise variable in the \\(Z\\) vector are labelled Z, while the variables in the \\(V\\) and \\(X\\) vectors are labelled V and X respectively."
plot_imps(imps[grepl('Interaction', scenario), ])
```

```{r var-imps}
cap <- "Mean variable importance across stages and model replicates. Importances are given as a percentage of the most important variable in each model. Standard deviations are shown in parentheses. Mean and standard deviation variable importance over all of the importances for each noise variable in the \\(Z\\) vector are labelled \\(Z\\), while the mean and standard deviation importance across all variables in the \\(V\\) and \\(X\\) vectors are labelled \\(V\\) and \\(X\\) respectively."
cap_toc <- "Mean variable importance across stages and model replicates"
col_heads <- c("B", "N", "PN", "I", "I+N", "I+PN")
var_levs <- c("M", "W", "D", "X", "Z", "V")
var_labs <- str_c("\\(", var_levs, "\\)")
imps %>%
  mutate(var_grp = factor(var_grp, levels = var_levs, labels = var_labs)) %>%
  group_by(scenario, mod, var_grp) %>%
  summarise(mean_imp = paste0(
    sprintf("%.1f", round(mean(importance), 1)), " (",
    sprintf("%.1f", round(sd(importance), 1)), ")")
  ) %>%
  spread(key = scenario, value = mean_imp) %>%
  ungroup() %>%
  select(-mod) %>% 
  knitr::kable(booktabs = T, col.names = c("", col_heads), caption = cap) %>% 
  group_rows("MARS", 1, 6) %>%
  group_rows("RF", 7, 12) %>%
  group_rows("CART", 13, 18)
```

Variable importances as a percentage of the most important variable for each method, stage, and training set replicate are shown in Figures \@ref(fig:var-imps-no-int) and \@ref(fig:var-imps-int). The interaction variables \(X_{1}\), \(X_{2}\), noise variables with no predictive value, \(V_{1}, ..., V_{90}\), and noise variables with some predictive value in the scenarios with predictive noise, \(Z_{1}, ..., Z_{10}\) are shown together as \(X\), \(V\), and \(Z\) respectively. Specifically, the Z label in Figures \@ref(fig:var-imps-no-int) and \@ref(fig:var-imps-int) shows violin plots of the relative importance for all of the variables in the \(Z\) vector for each scenario (and for every stage, so 3000 importances for each violin). Table \@ref(tab:var-imps) shows the mean and standard deviation of the relative importances over the 100 different training models and the three stages, for each variable group vector and each method. For example, the average importance of a \(V\) variable for all MARS models (at all stages) in the noise scenario (N) is 5.4, with a standard deviation of 11.1. Mean importances were quite similar between stages.

As shown in Figures \@ref(fig:var-imps-no-int) and \@ref(fig:var-imps-int) and Table \@ref(tab:var-imps), \(M\) and \(W\) are the most important variables across all model types. This aligns well with the higher importance of \(M\) and \(W\) on survival in the simulated data. Further, \(M\) is almost always the most important variable, which agrees with the higher weight that was placed on it in the simulation relative to \(W\).

Figures \@ref(fig:var-imps-no-int) and \@ref(fig:var-imps-int) and Table \@ref(tab:var-imps) also show that this version of MARS is much better than the other modeling techniques at identifying interacting variables and the treatment (\(D\)) as important. However, it is also much more prone to identify noise variables as important. As with mean survival times, there is higher variance across training sets in the importances MARS assigns to treatment and interacting variables.

Random forest is best at identifying the noise variables as not important, but also gives a relatively low importance to interacting and treatment variables. A single regression tree (CART) shows a similar trend in importances as the random forest models, but with near zero importance for noise variables and lower importance for dose and interacting (\(X\)) variables.
